### 2021-08-30 Manuscript/S3.Search_guide.Rmd
initial commit

---
title: "Databases' search guide"
date: "`r lubridate::today()`"
output:
  github_document:
    toc: yes
---

## WOS
- activate your institution proxy or log in in some way (it is only needed to get the API key, then you can log off the proxy)
- go to https://apps.webofknowledge.com/
- extract the API key from the URL, is denominated by SID
- put the key in the secrets.R file and source it or source Setup.R
- perform the search using the API

## IEEE
- get your API key and put it into the secrets.R
- source the file source Setup.R
- perform the search using the API

## Scopus
- change the NOT in the query to AND NOT
- go to https://www.scopus.com/search/form.uri?display=basic#basic
- log in with your institution credential
- set up the year range
- search within: title, abstract, keywords, authors
- select years in order to have just less than 2000 records
- download csv (Excel). Choose "Citation information" and "Abstract & keywords"
- call the files Scopus#.csv with # being a sequential number
- perform_search_session() will find and parse the files

## Pubmed
- put your API key in the secrets.R file and source it or source Setup.R
- perform the search using the API
- put the same query into pubmed and download records in the Pubmed format.
- only 10k records can be downloaded at once, so use the year filter to partition the records
- call the file Pubmed#.nbib with # being a sequential number
- perform_search_session() will find and parse the files

## Embase
- Login from within the institution or activate remote access
- go to https://www.embase.com/#advancedSearch/default
- set your date range
- in the mapping field select all apart from "Limit to terms indexed in article as 'major focus'"
- in "Sources" select Embase
- select records in blocks of 10K
- chose format: Csv - fields by column, output: Full record (may take a while to prepare the download and you can download only one batch at time)
- call the files Embase#.csv with # being a sequential number