### 2021-10-11 Manuscript/Manuscript_Discussion_TD.Rmd
update

---
title: "Discussion"
date: "`r lubridate::today()`"
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    includes:
      in_header: header.tex
  github_document: default
---

## Discussion

We propose a new integrated framework to help researchers collect and screen scientific publications characterised by high performance and versatility, joining the growing field of systematic review automation (SRA) and helpers (SRH) tools [@cohen2006reducing; @cohen2010evidence; @ananiadou2009supporting; @o2015using].
By sharing the tool as an open-source R package and following a modular design, we tried to adopt some of the so-called Vienna Principles advocated by the International Collaboration for the Automation of Systematic Reviews (ICASR) [@beller2018making].\
Our tool thus consists of three main components: 1) an integrated query-based citation search and management engine, 2) a text-to-feature extractor module, 3) an active machine learning-based citation screener, and 4) a data-driven search query generation algorithm.\

The framework's search engine module is capable of automatically collecting citation data from three well-known scientific databases (i.e. Pubmed, Web of Science, and the database of the Institute of Electrical and Electronics Engineers) as well as process manually downloaded results from both the mentioned and other (SCOPUS, EMBASE) databases.
In comparison, most SRH tools, commercial or free to use, rely either on internal databases (e.g., Mendeley https://www.mendeley.com/) sometimes focusing just on a particular topic [@visser2010performing] or on a single external data source [@thomas2007eppi; @poulter2008mscanner; @soto2019thalia].\
Mixing different databases is fundamental to have a more comprehensive view of the literature [@bajpai2011search; @wilkins2005embase; @woods1998medline]:
in our results, 18.7% of the positive matches found were unique for one of the different data sources, and no positive record was present in all of them (data not shown).\ 
The online search algorithm is efficient enough to manage tens of thousands of search results, using various expedients to overcome the limitations of the online database in terms of traffic and download quota.\

The feature extractor module uses modern NLP techniques [@ananiadou2006text; @cohen2008getting] to transform text into input data for machine learning.
We did not include classical n-grams [@schonlau2017text], but we used network analysis to find non-consecutive frequently associated terms, a generalisation of n-grams relaxing the term adjacency assumption.
A similar network approach was applied to merge redundant terms to make model estimation more efficient and reduce possible noise.
The use of concurrency network-driven modelling of text is not new [@rousseau2015graph; @violos2016sentiment; @rousseau2015text] and is a valuable tool to extract semantic information not evident in one-word or consecutive n-gram models.\

The framework's record screening module implements active machine learning-based classification [@settles2009active, @miwa2014reducing] based on Bayesian Additive Regression Trees (BART) [@chipman2010bart; @kapelner2013bartmachine], adopting best practices from other SRA studies and bringing improvements at various levels.\
As with other boosted trees algorithms [@hastie2009boosting], BART can explore complex non-linearities, perform variable selection, manage missing data while sporting high performance in predictive power.\
However, its Bayesian foundation provides further benefits:
less sensitivity on hyperparameter choices, natural regularisation through priors, and, especially, predictive distributions as output in place of point-wise predictions [@soria2011belm; @joo2020being; @jospin2020hands].
By selecting relatively tight prior distributions, we discouraged excessively deep trees, long sequences of trees, or extreme predicted probabilities, to decrease the risk of overfitting.\
The algorithm runs multiple replications of the model and averages their predictive distributions to create an "ensemble", a technique that has been shown to improve out-of-sample predictive performance [@zhou2021ensemble; @dietterich2000ensemble] as confirmed by the hyperparameter evaluation (Suppl. Mat. S2).
This approach reduces the uncertainty in the predictive distribution tails due to the randomness in the MCMC fit [@robert2004monte], generating a shift of probability mass towards the centre and stabilising the distributions (i.e., decreasing variance without impacting bias).
On the other hand, just imposing robust uninformative priors against extreme predictions would have decreased variance but also shifted the distribution towards a non-decision zone, increasing bias [@hansen2000bayesian].\
Since the number of model replications significantly impacts computation times, we decided to use ten replicas, the lower value after which performance stabilises.\
We also evaluated whether bootstrapping between replications [@breiman1996bagging] would improve performance, but, contrary to theory, it was demonstrated to be slightly detrimental (Suppl. Mat. S2) compared to same-data ensembling [@diez2015diversity].\

A low rate of relevant matches (class imbalance) is typical in literature reviews [@sampson2011precision; @wallace2010semi; @o2015using], and such strong imbalance between positive and negative records can affect sensitivity [@khoshgoftaar2010comparing; @chawla2004special].\
To overcome the problem, we oversampled [@batista2004study] the positive records ten times before model fitting.
Our hyperparameter analysis showed that together with model ensembling, the oversampling rate was the parameter with the highest impact on performance.\
A known risk with positive oversampling is the misclassification of negative records [@ramezankhani2016impact].
However, since all predicted positives get manually reviewed in our approach, we are always ensured to achieve 100% specificity/positive predictive value: the only price for the increased sensitivity due to oversampling is a larger number of records to review.\
An alternative to oversampling would be applying different weights and/or cost to the classes [@abd2013review; @diez2015diversity], but the BART implementation we used did not have this feature;
also, using simple oversampling permits broader compatibility with different modelling engines [@galar2011review; @roshan2020improvement].\
Finally, ordering the records by query term frequency (simple query ordering) generate a far higher rate of relevant records in the initial training set (17.2%) compared to the overall data (0.11%), and this boosts the sensitivity of the model.\

One of the central innovations we introduced is the concept of "uncertainty zone", which derives from the Bayesian foundation of the classification model.
This construct guides the selection of the records to review, dynamically updating and shrinking after every CR iteration, as more uncertain predictions are evaluated (Suppl. Mat. S2 Fig. 1).
This approach permits overcoming the limitations of active learning related to selecting a stopping rule which focus on hard thresholds or propose singular new items at each iteration instead of a batch [@laws2008stopping; @miwa2014reducing; @zhu2010confidence]. Our algorithm instead requires more general parameters, like the PPD intervals based on which the uncertainty zone is built and the maximum number of iterations with no positive matches after which a session is concluded; also, the hyperparamer evaluation shows that the algorithm is robust against variations in these parameters.\
Since researchers are asked to review both records with a positive predicted label and those inside the uncertainty zone, this method can be considered as a unifying synthesis of the "certainty" and "uncertainty" paradigms of active learning [@miwa2014reducing].\

We evaluated performance as the capability of the screening procedure (automatic classification plus manual review) to find the largest number of relevant records while reviewing as few of them as possible (i.e., sensitivity $\times$ efficiency).
We avoided the classic out-of-sample approaches like train-test sampling, out-of-bag bootstrapping or cross-validation [@kohavi1995study; @james2013introduction].
Such methods primarily assume that the probability of a positive match is equal on average in both the training and test set [@tashman2000out];
this uniformity is broken by how the initial training set and the subsequent reviewed records are selected by the query-based ordering and the active learning algorithm, determining a lower positivity rate in the unlabelled records (Fig. 2).
Also, a literature corpus is unique per search query/database combination, and therefore any out-of-sample performance estimate is not replicable since no new data can be acquired related to the current corpus.\
Instead, to estimate overall sensitivity, we implemented a surrogate, strongly regularised [@van2021bayesian], Bayesian regression model relying on just the lower interval of the generated PPD as predictor to abstract the active learning-based classification process and achieve a maximum entropy [@harremoes2001maximum] probabilistic estimate of the number of missed positive matches among the unreviewed records.
This simple surrogate model fitted the data very well ($R^2$ consistently above 97%), and its posterior predictive distribution could be exploited to explore worse case scenarios in terms of sensitivity.\

Our framework achieved very high sensitivity by screening a markedly small fraction of all records, bringing a sensible reduction in workload.\
Based on the surrogate model, we predicted a theoretical median sensitivity of 100% [93.5%, 100%] in the first session (screening 4.29% of records) and of 97.3% [73.8%, 100%] in the second (screening 1.34% of records):
efficiency increased significantly in the second session since only a few new positive matches were found, but given the large number of records, uncertainty regarding sensitivity also expectedly increased.\
Both results are above the usual performance in the field [@o2015using] and in line with the 92% average sensitivity estimated after human-only screening [@edwards2002identification].
In one interesting case, the model spotted a human misclassification in the initial training set, demonstrating its robustness and value as a second screener, as already suggested by previous studies [@frunza2010building; @bekhuis2012screening; @bekhuis2010towards].
Finally, albeit the simple query ordering already concentrated most of the relevant matches in the first 20-25 thousand records, without the tool support some relevant records would have required almost the complete data set to be manually checked to be found.\

The model took ~5-20 minutes per iteration to perform predictions in session 1 (17,755 documents) and 20-40 minutes in session 2 (98,371 documents) on an eight-core, 2.5 GHz, 16 GB RAM laptop from 2014;
including manual record review, one session required 1-3 days of work, for a total of 1-2 weeks for the whole process (including record collection).
That is a considerable saving of time compared to the months usually required for the screening phase of systematic reviews [@bannach2019machine; @borah2017analysis; @allen1999estimating].
To our knowledge, our data sets are larger than what is typical in most SRA studies [@o2015using; @olorisade2016critical], emphasising the reliability of the tool in real-world scenarios.\

The last module of our framework is a data-driven query generation algorithm.
Creating an efficient and efficacious search query is a complex task [@lefebvre2011searching; @hammerstrom2010searching] since it requires building a combination of positive and negative terms to maximise the number of relevant search results while minimising the total number of results.
Our solution joins a sensitivity-driven subquery proposal engine based on concurrent decision trees [@blanco2019machine; @moore2018transparent] built on the BART ensemble PPD with a human review step and an efficiency-driven query joiner.
The aim is to generate a second query that helps find records missed during the first session search.
The generated query allowed indeed to retrieve few more positive matches not found in session 1, but at the cost of a significant increase in the number of documents.\
One interesting aspect of this functionality is that it provides a human-readable overview of the classification rules learned by the classification model, showing which combination of terms was particularly relevant and even spotting authors and geographical locations associated with the study topic.
The generated query, therefore, acted as a tool for machine learning explainability [@bhatt2020machine; @burkart2021survey], a feature useful to understand and spot bias in black-box classification algorithms [@malhi2020explainable];
explainability is often required or even legally mandatory for high-stake machine learning applications [@bibal2021legal; @bibal2020impact].\
It is important to note that this process is entirely data-driven.
The algorithm is only aware of the "world" defined by the data set generated by a specific search query focused on a particular topic.
Therefore, the new query may not be specific once applied to an unbounded search domain, returning an unmanageable amount of unrelated results.
The solution we found was to add another component specifying the general topic (antimicrobial resistance and healthcare-associated infections) of our research to the query.\

As reported, our framework builds on modularity.
We designed it to easily implement complete independence of the main modules in future iterations, making it possible for users to add custom features like citation search and parsing for other scientific databases, alternative text processing algorithms or machine learning modules.
We deem such interoperability extremely relevant because the main strength of our tool is the composition of many solutions and the general approach related to Bayesian active machine learning.
However, each of its components could benefit considerably from the recent improvements in text mining.\
For example, our text processing approach is based on the boolean bag-of-words paradigm, and indeed it could be improved by more nuanced text representations.
It could be evaluated if feature transformations like TF-IDF [@baeza1999modern; @ananiadou2006text] would be advantageous, even if we hypothesise that tree-based classification algorithms like BART are robust enough not to need such operation.
Word embedding, a technique that transforms terms in semantic vectors derived from the surrounding text [@turian2010word; @bollegala2015embedding, @minaee2021deep], could be used to eliminate semantically redundant terms or differentiate identical terms with different meanings given the context.
Another option would be to employ unsupervised learning models like Latent Dirichlet Analysis or Latent Semantic Analysis [@pavlinek2017text; @chen2016short; @landauer1998introduction] to extract topics to enrich the feature space.\
Our classification algorithm can be implemented with any Bayesian supervised machine learning method that provides full PPDs;
therefore alternative models could be evaluated, like Gaussian Processes which are known for their flexibility [@jayashree2020evaluation; @chen2015gaussian].
Even more interesting would be to test advanced learning algorithms that surpass the bag-of-words approach, taking into consideration higher-level features in the text like term context and sequences, long-distance term relationships, semantic structures, etc., [@cheng2019document; @minaee2021deep; @li2020survey; @yang2020survey; @lai2015recurrent; @farkas1995document], given that a Bayesian implementation of such algorithms is available (for example @chen2018approximate).\

The field of literature review automatisation is becoming mature and ripe with innovation, and we expect an increasing use of such technologies to confront the ever-faster production of scientific literature.
We believe it is appreciable that a multiplicity of tools are being made available to let researchers and policymakers find the instrument that better fits their needs.
We tried to contribute to the field with an innovative methodology and tool that provides excellent performance and easy integration with existing systematic review pipelines.