### 2022-02-03 Manuscript/Manuscript.Rmd
uodate abstract

---
title: An open-source integrated framework for the automation of citation collection and screening in systematic reviews. 
authors:
  - name: Angelo D'Ambrosio
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: angelo.d.ambrosio@uniklinik-freiburg.de; a.dambrosioMD@gmail.com
  - name: Hajo Grundmann
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: hajo.grundmann@uniklinik-freiburg.de
  - name: Tjibbe Donker
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: tjibbe.donker@uniklinik-freiburg.de
abstract: |
  The exponential and prolific growth of published original scientific contributions makes secondary literature abridgements increasingly demanding.
  We introduce a new open-source framework for systematic reviews that significantly reduces time and human resource allocation for text identification, collection, and screening of scientific literature.\
  The framework introduces three main tools:
  1) an automatic citation search engine and manager that allows collecting records from multiple citation databases (Pubmed, WOS, SCOPUS, EMBASE, IEEE) with a unified query syntax;
  2) a citation screening tool based on Bayesian active machine learning and natural language processing, which requires users feedback only on uncertain classifications to iteratively increase predictive accuracy;
  3) a semi-automatic data-driven query generator to create new search queries from existing reviewed citation data sets.\
  The framework was tested on an example topic to perform citation collection and screening of related scientific production.
  To evaluate the performance of the active machine learning classifier,
  we estimated the median posterior Sensitivity and Efficiency [90% Credible Intervals] using Bayesian simulation to predict the distribution of potentially relevant matches among the records not manually reviewed.\
  17755 unique records were collected through the framework citation manager;
  101 over 766 records were found to be relevant after manual evaluation,
  while the rest were excluded by automatic classification;
  the expected Efficiency was 95.7% [95.3%, 95.7%] with a Sensitivity of 100% [93.5%, 100%].\
  An additional search query generated from the labelled dataset,
  retrieved 82579 additional records of which only 567 records required human review and six additional positive matches were retrieved.
  Including the additional records, the overall expected Sensitivity decreased to 97.3% [73.8%, 100%] while the Efficiency achieved was 98.6% [98.2%, 98.7%].\
  For large studies, the developed framework significantly reduces workload, human effort and error,
  simplifying citation collection and screening with exceptional sensitivity,
  hence improving the standardization and repeatability of systematic reviews.
keywords:
  - Systematic review automation
  - Citation management
  - Online data collection
  - Active machine learning
  - Natural language processing
  - Bayesian modeling
bibliography: references.bib
lang: en-GB
output: 
  rticles::arxiv_article:
    number_sections: no
    keep_tex: true
    includes:
      in_header: header.tex
---

```{r, echo=FALSE, results='asis', message = FALSE, warning = FALSE, error = FALSE, dpi = 400, fig.width=12, fig.height=5}

out <- unlist(lapply(c(
  "Manuscript_Introduction.Rmd",
  "Manuscript_Methods.Rmd",
  "Manuscript_Results.Rmd",
  "Manuscript_Discussion.Rmd"
), knitr::knit_child, quiet = TRUE))

cat(out, sep = "\n")
```

\newpage

# References