### 2022-02-19 Manuscript/Manuscript.Rmd
update for arxiv submission

---
title: An open-source integrated framework for the automation of citation collection and screening in systematic reviews. 
authors:
  - name: Angelo D'Ambrosio
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: angelo.d.ambrosio@uniklinik-freiburg.de; a.dambrosioMD@gmail.com
  - name: Hajo Grundmann
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: hajo.grundmann@uniklinik-freiburg.de
  - name: Tjibbe Donker
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: tjibbe.donker@uniklinik-freiburg.de
abstract: |
  The exponential and prolific growth of published original scientific contributions makes secondary literature abridgements increasingly demanding.
  We introduce a new open-source framework for systematic reviews that significantly reduces time and human resource allocation for text identification, collection, and screening of scientific literature.\
  The framework provides three main tools:
  1) an automatic citation search engine and manager that allows collecting records from multiple citation databases (Pubmed, WOS, SCOPUS, EMBASE, IEEE) with a unified query syntax,
  2) a citation screening tool based on Bayesian active machine learning and natural language processing that requires users feedback only on uncertain classifications to iteratively increase predictive accuracy and,
  3) a semi-automatic, data-driven query generator to create new search queries from existing reviewed citation data sets.\
  The framework was tested on an example topic to collect and screen citations of related scientific production.
  To evaluate the performance of the active machine learning classifier,
  we estimated the median posterior sensitivity and efficiency [90% Credible Intervals] using Bayesian simulation to predict the distribution of potentially relevant matches among the records not manually reviewed.\
  A total of 17,755 unique records were collected through the framework citation manager;
  101 of 766 records were found to be relevant after manual evaluation,
  while the rest were excluded by automatic classification;
  the theoretical efficiency was 95.6% [95.3%, 95.7%] with a sensitivity of 100% [93.5%, 100%].\
  A new search query was generated from the labelled dataset, and 82,579 additional records were collected;
  only 567 records required human review after automatic screening, and six additional positive matches were found.
  Including the additional records, the overall expected sensitivity decreased to 97.3% [73.8%, 100%] while the efficiency increased to 98.6% [98.2%, 98.7%].\
  For large studies, the framework can significantly reduce the human resources required to conduct systematic reviews
  by simplifying citation collection and screening while demonstrating exceptional sensitivity.
  Such a tool can improve the standardization and repeatability of systematic reviews.
keywords:
  - Systematic review automation
  - Citation management
  - Online data collection
  - Active machine learning
  - Natural language processing
  - Bayesian modeling
bibliography: references.bib
lang: en-GB
output: 
  rticles::arxiv_article:
    number_sections: no
    keep_tex: true
    includes:
      in_header: header.tex
---

```{r, echo=FALSE, results='asis', message = FALSE, warning = FALSE, error = FALSE, dpi = 400, fig.width=12, fig.height=5}

out <- unlist(lapply(c(
  "Manuscript_Introduction.Rmd",
  "Manuscript_Methods.Rmd",
  "Manuscript_Results.Rmd",
  "Manuscript_Discussion.Rmd"
), knitr::knit_child, quiet = TRUE))

cat(out, sep = "\n")
```

\newpage

# References