### 2022-02-19 Manuscript/S1.Protocol.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\title{Study protocol}
\author{}
\date{\vspace{-2.5em}2022-02-19}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Study protocol},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\newpage

\hypertarget{research-protocol}{%
\section{Research Protocol}\label{research-protocol}}

This document will guide the reader along the steps to utilise the
framework and reproduce our results.

Add installation and setup info here.

\hypertarget{acquisition-of-the-citaion-data}{%
\subsection{Acquisition of the citaion
data}\label{acquisition-of-the-citaion-data}}

Once the framework is loaded, the user defines an initial search query
which needs to be as specific as possible while still generating a
sufficient number of positive matches:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{source}\NormalTok{(}\StringTok{"R/Setup.R"}\NormalTok{) }\CommentTok{\# Load the framework}

\CommentTok{\# Initial query to be built on domain knowledge. It accepts OR, AND, NOT boolean}
\CommentTok{\# operators and round brackets to group terms.}
\NormalTok{query }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}((model OR models OR modeling OR network OR networks) AND}
\StringTok{(dissemination OR transmission OR spread OR diffusion) AND (nosocomial OR}
\StringTok{hospital OR "long{-}term{-}care" OR "long term care" OR "longterm care" OR}
\StringTok{"long{-}term care" OR "healthcare associated") AND (infection OR resistance OR}
\StringTok{resistant))\textquotesingle{}}

\CommentTok{\# Year filter. The framework converts it to the API{-}specific format seamlessly.}
\CommentTok{\# common logical comparators can be used, i.e. \textless{}, \textless{}=, \textgreater{}, \textgreater{}=, while dashes}
\CommentTok{\# denotes inclusive date intervals. A single year restricts results to one year}
\CommentTok{\# period.}
\NormalTok{year\_filter }\OtherTok{\textless{}{-}} \StringTok{"2010{-}2020"}
\end{Highlighting}
\end{Shaded}

The query is passed to the \texttt{perform\_search\_session()} function,
together with an unique identifier for the \textbf{search session} and
\textbf{search query}. A session identifies a homogeneous group of
complementary or alternative queries (e.g.~sometimes it is easier to
create two different simpler queries than a complex and long one that
returns the same results).\\
In our implementation, we define as sessions the group of actions that
includes the utilisation of a search query and the subsequent iterations
of manual review and automatic classification of its results.\\

\texttt{perform\_search\_session()} will use common scientific research
database APIs (at the moment Pubmed/Medline, Web of Science and IEEE) to
acquire records related to the query. The records will be stored as CSV
files in folders with the following structure:
\texttt{Records/session\_id/query\_id}, created automatically. The
function also creates or updates a \textbf{journal} (a CSV or Microsoft
Excel file) with all the information about sessions, queries and search
results.\\
Some services (IEEE, WOS) requires an API key to access the data; the
API key is not mandatory for Pubmed but can help avoid quota limitations
in case of a large number of results. IEEE can also be searched without
API: the framework will utilise a web scraping approach if the API key
is missing, but the success of this method is not guaranteed. See
Supplemental Material S3 for detailed instruction on preliminary steps
and tips for each supported search engine. These keys can be stored in
the \texttt{secrets.R} file in the working directory, with the following
example code (the ``\emph{baysren}'' prefix identifies the package
specific options):

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{options}\NormalTok{(}\AttributeTok{baysren.ieee\_api\_key =} \StringTok{"your{-}ieee{-}api{-}key"}\NormalTok{)}
\FunctionTok{options}\NormalTok{(}\AttributeTok{baysren.ncbi\_api\_key =} \StringTok{"your{-}ncbi{-}api{-}key"}\NormalTok{)}
\FunctionTok{options}\NormalTok{(}\AttributeTok{baysren.wos\_api\_key =} \StringTok{"your{-}wos{-}api{-}key"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This file will be automatically parsed from the framework.

If previously downloaded results are already available, users can create
the \texttt{Record} folder manually and put the data there (the file
names need to contain the source name as written in the \texttt{sources}
argument of \texttt{perform\_search\_session()}). The function will
acquire and parse them into a standard format, even if split into
multiple files due to download limits (e.g.~\texttt{Pubmed1.nbib},
\texttt{Pubmed2.nib}, etc.). In addition to Pubmed, WOS and IEEE, this
method permits to import also EMBASE and SCOPUS records, for which an
API search was not yet implemented.\\
Using the \texttt{actions} argument, users can choose whether to perform
an API search, parse already downloaded results, or both.

Note that for Pubmed, the API search may return slightly fewer records
than a manual one due to a different query expansion algorithm (this
divergence should be solved in April 2022) between the two services,
therefore is advisable to perform also a manual search and put the
resulting .nbib files in the \texttt{Session/Query} folder.
\texttt{perform\_search\_session()} will acquire them seamlessly.

If a user wants to perform the API searches and results parsing manually
for each source, the framework exposes the individual functions to
search for records using the APIs and parsing bibliography files.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# This call will perform an API search on Pubmed, WOS and IEEE services and/or}
\CommentTok{\# (based on the \textasciigrave{}actions\textasciigrave{} argument) parse already downloaded data (in our case,}
\CommentTok{\# additional Pubmed results and Embase and Scopus results) manually added to the}
\CommentTok{\# Session1/Query1 folder.}

\NormalTok{journal }\OtherTok{\textless{}{-}} \FunctionTok{perform\_search\_session}\NormalTok{(}
  \AttributeTok{query =}\NormalTok{ query, }\AttributeTok{year\_query =}\NormalTok{ year\_filter,}
  \AttributeTok{session\_name =} \StringTok{"Session1"}\NormalTok{, }\AttributeTok{query\_name =} \StringTok{"Query1"}\NormalTok{,}
  \AttributeTok{records\_folder =} \StringTok{"Records"}\NormalTok{,}
  \AttributeTok{journal =} \StringTok{"Session\_journal.csv"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Once the records are stored, they must be read and merged into an
\textbf{annotation file}, where the initial manual review will be
performed.\\
A series of functions are available to prepare the data for manual
evaluation:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Extract the file paths of records. Arguments can be used to filter by session}
\CommentTok{\# query, source. Only parsed files will be returned, not the raw ones downloaded}
\CommentTok{\# manually.}
\NormalTok{Annotation\_data }\OtherTok{\textless{}{-}} \FunctionTok{extract\_source\_file\_paths}\NormalTok{(journal) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Read record files. Will parse them if raw data is downloaded manually (not}
  \CommentTok{\# necessary in this case). Return a list of records, one per file.}
  \FunctionTok{read\_bib\_files}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Join a list of records into one dataframe, solving duplicated records}
  \FunctionTok{join\_records}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Order records by the relative frequency of the query terms in the title +}
  \CommentTok{\# abstract text. Increases the chance of encountering relevant records at the}
  \CommentTok{\# beginning of the manual classification.}
  \FunctionTok{order\_by\_query\_match}\NormalTok{(query) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Add fields for the manual classification}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{Rev\_manual =} \ConstantTok{NA}\NormalTok{, }\CommentTok{\# Where the first manual classification will be made}
    \AttributeTok{Rev\_prediction =} \ConstantTok{NA}\NormalTok{, }\CommentTok{\# Column for the evaluation of the predicted classes}
    \AttributeTok{.before =}\NormalTok{ DOI}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

All these steps are joined into the \texttt{create\_annotation\_file()}
function. This function can take as input either the path to record
files, or the parsed records (one data frame or a list of data frames)
or a group of folders where to look for files into./ In addition, it
also allows adding the new records to a previous annotation file (arg
\texttt{prev\_records}). This is useful when performing consecutive
search sessions with different queries; previous manual classifications
can be imported using \texttt{prev\_classification}.\\
If the original query is passed to the \texttt{reorder\_query} argument,
the records are reordered according to the frequency of query terms in
the title and abstract (\textbf{simple query ordering}); this approach
lets increase the rate of relevant matches at the beginning of the
dataset.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# This function extracts the appropriate file paths from a session journal. By}
\CommentTok{\# default it passes all stored files, otherwise it can be filtered by session,}
\CommentTok{\# query, and source (i.e. Pubmed, WOS, IEEE)}
\NormalTok{record\_files }\OtherTok{\textless{}{-}} \FunctionTok{extract\_source\_file\_paths}\NormalTok{(journal)}

\CommentTok{\# create\_annotation\_file() accept a great variety of inputs:}

\CommentTok{\# either record file paths}
\NormalTok{input }\OtherTok{\textless{}{-}}\NormalTok{ record\_files}

\CommentTok{\# or specific record file folders}
\NormalTok{input }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(}\StringTok{"Records"}\NormalTok{, }\StringTok{"Session1"}\NormalTok{, }\StringTok{"Query1"}\NormalTok{)}

\CommentTok{\# or parent folders, since it searches for files recursively}
\NormalTok{input }\OtherTok{\textless{}{-}} \StringTok{"Records"}

\CommentTok{\# or the already parsed files}
\NormalTok{input }\OtherTok{\textless{}{-}} \FunctionTok{read\_bib\_files}\NormalTok{(record\_files)}

\CommentTok{\# We can then call create\_annotation\_file() with one of the above input}
\NormalTok{Annotation\_data }\OtherTok{\textless{}{-}} \FunctionTok{create\_annotation\_file}\NormalTok{(input, }\AttributeTok{reorder\_query =}\NormalTok{ query)}
\end{Highlighting}
\end{Shaded}

Once created, the annotations need to be stored in an \textbf{Annotation
Session} folder, which will contain all information necessary for the
automatic classification. This operation is streamlined by the
\texttt{create\_session()} function, which also helps manage duplicated
sessions.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create the first session}
\FunctionTok{create\_session}\NormalTok{(Annotation\_data, }\AttributeTok{session\_name =} \StringTok{"Session1"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The file annotation file is called \texttt{Records\_\{date\}}, situated
in the folder \texttt{Sessions/\{session\_name\}}; \texttt{date} here
represents the session creation timestamp, and \texttt{session\_name}
the name passed by the user.

Notice that the framework is ``pipe'' friendly, and the precedent steps
can be all performed in one call:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# First, create the Records/Session1/Query1 folder and put the manually}
\CommentTok{\# downloaded records there if there are any. Otherwise, the folders will be}
\CommentTok{\# created automatically during the API search}

\CommentTok{\# perform searches and parse files; save and return the journal}
\NormalTok{Records }\OtherTok{\textless{}{-}} \FunctionTok{perform\_search\_session}\NormalTok{(}
  \AttributeTok{query =}\NormalTok{ query, }\AttributeTok{year\_query =}\NormalTok{ year\_filter,}
  \AttributeTok{session\_name =} \StringTok{"Session1"}\NormalTok{, }\AttributeTok{query\_name =} \StringTok{"Query1"}\NormalTok{,}
  \AttributeTok{journal =} \StringTok{"Session\_journal.csv"}
\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# return the record file paths}
  \FunctionTok{extract\_source\_file\_paths}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# read, join, reorder records and save them in the Sessions folder}
  \FunctionTok{create\_annotation\_file}\NormalTok{(}\AttributeTok{reorder\_query =}\NormalTok{ query) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# Create a new session with the records}
  \FunctionTok{create\_session}\NormalTok{(}\AttributeTok{session\_name =} \StringTok{"Session1"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{manual-classification-and-prediction}{%
\subsection{Manual classification and
prediction}\label{manual-classification-and-prediction}}

Now the annotation file is ready for the manual classification of
records. The annotation file has a column called \texttt{Rev\_manual},
where records need to be classified as ``positive'' (y) or ``negative''
(n) matches by the user. We suggest manually labelling an
\textbf{initial training set} of 250 records, which should be sufficient
if a minimum \textasciitilde20\% positivity rate is present (see the
section about model hyperparameter optimisation)./ The function
\texttt{plot\_classification\_trends()} can help visualise the trend of
the positives/negatives as the number of classified records increases./

Once the initial training set is prepared, the function
\texttt{enrich\_annotation\_file()} will use a Bayesian Additive
Regression Trees (BART) machine learning model to learn the labelling
pattern given the record data and predict the positive match probability
for the unlabelled records.\\
\texttt{enrich\_annotation\_file()} comprises two steps. The first step
is the creation of a \textbf{Document Term Matrix} (DTM):

\begin{itemize}
\tightlist
\item
  The framework extracts \textbf{features} from text data (i.e., title
  and abstract) of the records, authors, keywords and MESH terms (see
  Methods for more info). Before features extraction, text data terms
  are lemmatised, stop-words are removed, and only nouns, adjectives,
  verbs and untagged terms are retained. Authors, keywords, and MESH are
  kept unmodified. Co-occurrent features (non-consecutive ngrams,
  \textbf{nc-grams}) are identified and \textbf{redundant features}
  (i.e.~terms always appearing together) are joined to decrease the
  feature space noise.
\item
  For each term, a 0/1 label specifies if it is present in each record.
\item
  Positive records are oversampled \textbf{10x} to improve sensitivity;
\item
  An extra feature is added for each record field (title, abstract,
  authors, keywords, MESH terms), reporting the number of terms in each.
\end{itemize}

The second step is the modelling of the labelling pattern and the
predictions of the relevancy/positivity probability:

\begin{itemize}
\tightlist
\item
  The function models the labelling pattern using the manually labelled
  records (**learning phaseéé);
\item
  After the model is created, it assigns a \textbf{posterior predictive
  distribution} (PPD) of the probability of positivity to each record
  (\textbf{prediction phase}). If the number of records is large, the
  prediction step is broken into chunks of \textbf{5000 records}, to
  avoid memory saturation; the users can set up this parameter according
  to their machine capabilities;
\item
  This process is repeated \textbf{10 times}, and the predictive
  distributions are averaged to induce shrinkage of the estimates (i.e.,
  decrease the chance of outliers), creating a prediction
  \textbf{ensemble};
\item
  The resulting distributions are summarised by the \textbf{98\%
  interval} {[}98\% PrI{]} and the median; the {[}98\% PrI{]} identifies
  the \textbf{uncertainty zone};
\item
  The unlabelled records will receive a positive/negative predicted
  label; the label is positive if its lower {[}98\% PrI{]} bound is
  higher than the uncertainty zone upper bound and negative if the upper
  {[}98\% PrI{]} bound is lower than the uncertainty zone lower bound.
  All other records are labelled as \textbf{uncertain} (unk);
\item
  If a predicted label does not match an existing one, the record will
  be labelled as \textbf{check}.
\end{itemize}

The user can define parameters like the positives' oversampling rate,
the number of models to average in the ensemble, whether to apply
bootstrap resampling before each repetition, the prediction chunk size,
and the PPD quantiles used to build the uncertainty zone. To use the
function, the only input needed is the session name; the function will
automatically identify the relevant file to make predictions on.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{New\_annotations }\OtherTok{\textless{}{-}} \FunctionTok{enrich\_annotation\_file}\NormalTok{(}\StringTok{"Session1"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This function will produce a new file which is stored in the Annotations
folder of the session; the file will contain a column called
\texttt{Rev\_prediction\_new} where the user will find records that
require manual review marked by ``*``. These are records with uncertain
labelling (''unk''), but also positive predictions and ``y''/``n''
predictions that contradict previous manually input labels (labelled as
``check''). This approach increases the accuracy of the predictions
since the machine learning system is set up to favour sensitivity at the
expense of specificity. If the function finds prelabelled records
(\texttt{Rev\_previous} column) added when the Annotation file was
created or passed as a further input to
\texttt{enrich\_annotation\_file()} (\texttt{prev\_classification}
argument), they will be used to label the records marked in the
\texttt{Rev\_prediction\_new} column.

The function adds a number of diagnostic pages to the annotation file:

\begin{itemize}
\tightlist
\item
  A results summary page reporting the number of records in need of
  manual review (which includes ``unk'', ``check'', and unreviewed ``y''
  records), the changes, and the final distribution of the labels. It
  also shows other information like the number of features in the DTM,
  the path to the parent file from which the annotation file was
  created, the replication number (see later). The content of this
  summary is also saved as a CSV file in the Results folder.
\item
  A variable importance page that lists the used terms in order of the
  fraction of BART trees in which they were used. If an ensemble of
  models is used, a more robust score is computed, taking the ratio of
  the mean of the term tree fraction among models and its standard
  deviation.
\item
  The arguments passed to the function if different from the default,
  useful for reproducibility.
\item
  A classification performance summary. The procedure may take some time
  and requires \texttt{cmdstan} installed on the machine. Therefore the
  \texttt{compute\_performance} argument is set to \texttt{FALSE} by
  default.
\end{itemize}

Finally, the PPD ensemble produced by the Bayesian engine is stored on
the disk. This feature can be turned off to save disk space, but at
least the samples of the last annotation are necessary to build a new
search query (see next section).

The process of predicting labels and manually reviewing them is what
defines a ``classification-review (CR) iteration''.

Once the manual review of the generated annotation file is completed,
the user can \textbf{rerun exactly the same function}. The last reviewed
annotation file will be automatically picked up and used for a new CR
iteration. If the number of records requiring manual review is
excessively high (e.g.~\textgreater{} 250), the function needs to be
called with the \texttt{stop\_on\_unreviewed} argument set to
\texttt{FALSE} to allow the classification to proceed even in the
presence of unreviewed records in the \texttt{Rev\_prediction\_new}
column.

If there are no positive results (after manual review) in a CR
iteration, the framework records the next iteration as a
\textbf{replication}. After four replications with no new positive
matches, the function will prevent another iteration and send a warning
message, indicating that the session is complete. The user can change
the replication limit; furthermore, it is possible to prevent the
function from starting if either a specified number (or a fraction) of
the total records has been manually reviewed or if a certain number of
positive matches have been found. These three parameters are passed to
the \texttt{limits} argument, which has the following structure and
defaults:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{list}\NormalTok{(}
  \AttributeTok{stop\_after =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{pos\_target =} \ConstantTok{NULL}\NormalTok{,}
  \AttributeTok{labeling\_limit =} \ConstantTok{NULL}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The result files in the Results folder are generated after the
Classification step but before the Review one, and therefore show
preliminary results with still unreviewed records. The function
\texttt{consolidate\_results(session\_name)} will update the results
files to show the results after the manual review. The original result
data is stored anyway in the respective annotations files.

\hypertarget{new-search-query-generation}{%
\subsection{New search query
generation}\label{new-search-query-generation}}

It is not uncommon for the first query used in a systematic review not
to be the most effective one, especially in terms of sensitivity. We
tried to address the problem by creating an automatic query generation
system.\\
The solution is experimental and greatly favours sensitivity at the
expense of specificity. The general query generation mechanism is the
following:

\begin{itemize}
\tightlist
\item
  We fit a partition tree between the DTM and 800 samples from the PPD;
  if a term is present multiple times in the DTM (e.g., both title and
  abstract), they are counted just one, and field term count features
  are removed. This step generates a list of rules composed by
  \(AND\)/\(NOT\) ``conditions'' made of terms/authors/keywords/MESH
  tokens, which together identify a group of records.
\item
  For each rule, negative conditions (i.e., \(NOT\) statements) are
  added iteratively, starting from the most specific one, until no
  conditions are found that would not also remove positive records.
\item
  The extended set of rules is sorted by positive-negative record
  difference in descending order. The cumulative number of unique
  positive records is computed and used to group the rules. Rules inside
  each group are ordered by specificity.
\item
  The researcher is then asked to review the rule groups, selecting one
  or more rules (useful if they convey different meaning) from each, or
  edit them (in case too specific positive or negative conditions were
  included). It is possible to exclude a group of rules altogether,
  especially those with the worse sensitivity/specificity ratio.
\item
  The selected rules are joined together by \(OR\) statements, defining
  a subset of records with a sensibly higher proportion of positive
  records than the original one.
\item
  Redundant rules (i.e., rules whose positive records are already
  included in more specific ones) and conditions (i.e., conditions that
  once removed do not decrease the total number of positive or do not
  increase the negative records) are removed.
\item
  Finally, the rules are re-elaborated in a query usable on the major
  scientific databases.
\end{itemize}

The steps for generating a query are the following:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# The last annotation file and posterior sample matrix in the session are found;}
\CommentTok{\# the most predictive terms (based on the importance score) are chosen and}
\CommentTok{\# simple prediction trees are used to map the terms to a random subset of the}
\CommentTok{\# posterior probabilities (extracted from the posterior sample matrix saved}
\CommentTok{\# during the last CR iteration). The score threshold and the number of trees can}
\CommentTok{\# be chosen by the user; less stringent values can significantly increase}
\CommentTok{\# computation time. By default results will be saved in a file called}
\CommentTok{\# rule\_data.rds in the session folder.}

\NormalTok{candidate\_queries }\OtherTok{\textless{}{-}} \FunctionTok{extract\_rules}\NormalTok{(}\StringTok{"Session1"}\NormalTok{)}

\CommentTok{\# The \textquotesingle{}candidate\_queries\textquotesingle{} contains the rules, the DMT and the labels on which}
\CommentTok{\# the rules were built, which must be passed to generate\_rule\_selection\_set().}
\CommentTok{\# This function selects the most effective (higher positive/negative matches}
\CommentTok{\# delta) rules; also adds negative terms to increase specificity. A data frame}
\CommentTok{\# is generated (or automatically saved if save\_path is set). The output proposes}
\CommentTok{\# groups of equivalent (in terms of the number of positive matches found) groups}
\CommentTok{\# of query, ordered by specificity; for each group, the user needs to select the}
\CommentTok{\# best rule by setting the selected\_rule column to TRUE. The groups are in order}
\CommentTok{\# of growing sensitivity/decreasing specificity, so it may make sense to exclude}
\CommentTok{\# the last groups totally in order to avoid too many search results.}

\NormalTok{Target }\OtherTok{\textless{}{-}}\NormalTok{ candidate\_queries}\SpecialCharTok{$}\NormalTok{DTM}\SpecialCharTok{$}\NormalTok{Target}
\NormalTok{SpecificDTM }\OtherTok{\textless{}{-}}\NormalTok{ candidate\_queries}\SpecialCharTok{$}\NormalTok{SpecificDTM}

\CommentTok{\# No default is given for the save file, since generate\_rule\_selection\_set()}
\CommentTok{\# have no info about the session it\textquotesingle{}s working on.}
\NormalTok{selection\_set\_file }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(}\StringTok{"Sessions"}\NormalTok{, }\StringTok{"Session1"}\NormalTok{, }\StringTok{"Selected\_rules.xlsx"}\NormalTok{)}

\NormalTok{selection\_set }\OtherTok{\textless{}{-}} \FunctionTok{generate\_rule\_selection\_set}\NormalTok{(}
\NormalTok{  candidate\_queries}\SpecialCharTok{$}\NormalTok{rule,}
  \AttributeTok{target\_vec =}\NormalTok{ Target,}
  \AttributeTok{target\_data =}\NormalTok{ SpecificDTM,}
  \AttributeTok{save\_path =}\NormalTok{ selection\_set\_file}
\NormalTok{)}

\CommentTok{\# Once the manual selection is performed the selected rules can be further}
\CommentTok{\# simplified in order to remove redundant rules and terms. The aim is to provide}
\CommentTok{\# the shortest rule without loss of sensitivity.}

\NormalTok{simplified\_rules }\OtherTok{\textless{}{-}} \FunctionTok{file.path}\NormalTok{(}\StringTok{"Sessions"}\NormalTok{, }\StringTok{"Session1"}\NormalTok{, }\StringTok{"Selected\_rules.xlsx"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{import\_data}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{simplify\_ruleset}\NormalTok{(}\AttributeTok{target\_vec =}\NormalTok{ Target, }\AttributeTok{target\_data =}\NormalTok{ SpecificDTM) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pull}\NormalTok{(rule)}

\CommentTok{\# Finally, the selected rules need to be joined in a query that scientific}
\CommentTok{\# search engines can understand}

\NormalTok{new\_query }\OtherTok{\textless{}{-}} \FunctionTok{rules\_to\_query}\NormalTok{(simplified\_rules)}
\end{Highlighting}
\end{Shaded}

One last thing to keep in mind is that the query generation algorithm is
only aware of the literature collected inside the annotation file. So it
will only be able to pick term combinations that identify positive
matches inside the already specific set of documents while not
necessarily being discriminant from the rest of the global literature.\\
Once the final query is generated, it is advisable to add a constraining
subquery with an AND logic to restrict the results to the scope of the
topic of interest.\\
In our case, we added ``AND ((antimicrobial resistance) OR (healthcare
infection))'' since, being the global topic of our research, those terms
were understandably not selected by the query generation algorithm as
discriminant.\\

The final query was the following:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{query }\OtherTok{\textless{}{-}} \StringTok{"(((Donker T) NOT (bacterium isolate)) OR ((network patient) AND}
\StringTok{(resistant staphylococcus aureus) NOT (monte carlo) NOT isolation) OR}
\StringTok{(facility AND (network patient) AND regional NOT hospitals NOT increase NOT}
\StringTok{(patient transport) NOT (control infection use)) OR ((patient transport) NOT}
\StringTok{(Donker T) NOT worker) OR (hospitals AND (network patient) NOT}
\StringTok{(patient transport) NOT regional NOT clinical) OR (facility AND}
\StringTok{(network patient) NOT hospitals NOT (patient transport) NOT regional NOT}
\StringTok{prevention NOT medical) OR ((healthcare facility) NOT (Donker T) NOT}
\StringTok{worker NOT positive) OR (hospitals NOT (network patient) NOT medical NOT}
\StringTok{environmental NOT outcome NOT global) OR ((network patient) NOT facility NOT}
\StringTok{hospitals NOT (patient transport) NOT therapy NOT global)) AND}
\StringTok{((antimicrobial resistance) OR (healthcare infection))"}
\end{Highlighting}
\end{Shaded}

Another advantage of this technique is an at-glance view of relevant
concepts for the topic of interest.

\hypertarget{new-search-session}{%
\subsection{New search session}\label{new-search-session}}

Once the new query was generated, a second search session can be
created, using the strategy described in the protocol. Due to the lower
specificity of the second query, particular care may be needed in
managing the far larger amount of records: multiple record files may
need to be downloaded from Pubmed, EMBASE, and SCOPUS since they have
limits on the number of records one can download at once; also, the
machine learning classification processing time increases proportionally
with the number of records to label.\\
It is advisable to consider removing the less specific rules iteratively
from the new query until a manageable number of records is obtained,
considering that the probability of finding new positive matches not
already found by more specific queries drops exponentially.\\

Once the new records are collected are manually collected for databases
without an API access, a new search can be performed. The following code
describe how to create a new session. Keep in mind to give the session a
new name and to import the annotated records from the last session when
generating the new annotation file.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Repeat the search with the new query, giving a different Session name not to}
\CommentTok{\# override previous records}
\NormalTok{journal }\OtherTok{\textless{}{-}} \FunctionTok{perform\_search\_session}\NormalTok{(}
  \AttributeTok{query =}\NormalTok{ query, }\AttributeTok{year\_query =}\NormalTok{ year\_filter,}
  \AttributeTok{session\_name =} \StringTok{"Session2"}\NormalTok{, }\AttributeTok{query\_name =} \StringTok{"Query1"}\NormalTok{,}
  \AttributeTok{records\_folder =} \StringTok{"Records"}\NormalTok{,}
  \AttributeTok{journal =} \StringTok{"Session\_journal.csv"}
\NormalTok{)}

\CommentTok{\# Get the path to the annotated data in the previous session}
\NormalTok{previousAnnotations }\OtherTok{\textless{}{-}} \FunctionTok{get\_session\_files}\NormalTok{(}\StringTok{"Session1"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{Annotations }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{last}\NormalTok{()}

\CommentTok{\# Create the new annotation file passing the folder with the new records}
\NormalTok{Annotation\_data }\OtherTok{\textless{}{-}} \FunctionTok{create\_annotation\_file}\NormalTok{(}
  \FunctionTok{file.path}\NormalTok{(}\StringTok{"Records"}\NormalTok{, }\StringTok{"Session2"}\NormalTok{),}
  \AttributeTok{reorder\_query =}\NormalTok{ query,}
  \AttributeTok{prev\_records =}\NormalTok{ previousAnnotations}
\NormalTok{)}

\CommentTok{\# Create the new annotation session}
\FunctionTok{create\_session}\NormalTok{(Annotation\_data, }\AttributeTok{session\_name =} \StringTok{"Session2"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Once the new annotation session data is created,
\texttt{enrich\_annotation\_file(\textquotesingle{}Session2\textquotesingle{})}
can be used to perform the CR iterations. It may be possible that during
the first CR iteration, the number of records requiring manual review is
very large: we suggest evaluating still not more than 250, setting the
\texttt{stop\_on\_unreviewed} argument to \texttt{FALSE} to proceed in
the automatic classification even if some uncertain records from the
precedent iteration are still unreviewed.

\hypertarget{result-analysis-and-reporting}{%
\subsection{Result analysis and
reporting}\label{result-analysis-and-reporting}}

Several tools are available to evaluate the classification
performance.\\
As said before, a result file is stored in the session folder for each
CR iteration with information on the classification status; remember to
run \texttt{consolidate\_results(session\_name)} if you want these
results to account for the manual review of the predicted labels.\\
\texttt{compute\_changes()} is a simple function that take an annotation
data frame and computes the changes from the last annotation
iteration.\\
The primary function for result analysis is
\texttt{estimate\_performance()}. The function uses a Bayesian model to
estimate how the probability of a positive match drops as records' lower
boundary of the posterior predictive interval decreases. This model
computes the cumulative distribution of possibly missed positive
records, which permits extraction of the engine's theoretical
sensitivity and efficiency to lower 90\% credibility boundaries (i.e.,
the indicators are better than these values with 95\% probability).
These boundaries depict worse case scenarios since it is impossible to
know the actual number of missed positive records.\\
The function returns the Bayesian \(R^2\) of the model (values below
85-90\% would invalidate the model analysis), the posterior sensitivity,
the observed and theoretical number positives, the expected efficiency
given the theoretical number of positives (that is 1 - the number of
reviewed records above the number of records to review to find the
predicted number of positives according to a negative hypergeometrical
distribution). Also, the function plots the cumulative distribution of
the observed and theoretical number of positive records.\\
\texttt{extract\_var\_imp()} extract the most relevant terms used by the
BART model and summarises their relevancy for the machine learning model
and their statistical relevancy in a simple generalised linear model. A
discrepancy between these two scores indicates a non-linear
effect/interaction between terms.

Finally, several functions are available (check \texttt{R/Reporting.R})
to summarise and format the results.\\
For example, \texttt{summarise\_by\_source()} and
\texttt{summarise\_sources\_by\_session()}, respectively for one session
or a group of sessions, report the scientific databases the records are
collected from, while \texttt{get\_source\_distribution()} describes the
distribution of the number of common sources per record.\\
\texttt{summarise\_annotations()} and
\texttt{summarise\_annotations\_by\_session()} describe the results of
the CR iterations, again for one or multiple sessions, reporting the
number of positive and negative matches, the number of terms (features)
used by the machine learning model, and how the labels change between
iterations. \texttt{format\_performance()} and
\texttt{format\_var\_imp()} format the output of the similarly named
analyses. \texttt{plot\_predictive\_densities()} plots how the mixture
of posterior predictive distributions of records, divided into positive,
negative and to-review records, change between iterations.

\hypertarget{model-parameters-optimisation}{%
\section{Model parameters'
optimisation}\label{model-parameters-optimisation}}

As with many complex machine learning models, the choice of
hyperparameters can significantly influence the framework performance.
For a systematic review helper algorithm, we define the performance as
the ability to find as many positive matches by manually
labelling/reviewing as few records as possible:\\

\[
Sensitivity \times Efficiency = \\
\frac{Pos.\ found}{Tot.\ pos.} \times \bigg(1 - \frac{Records\ reviewed}{Tot.\ records}\bigg)
\]

We utilise a ``grid search'' algorithm to find the best set of
hyperparameters. We tested these tools on our dataset, manually
labelling the first 1200 records, ordered by simple query match.\\
The first step is to create a folder to store the file required for the
search, for example, \emph{Grid Search}. Then a fully reviewed record
file in the format generated by \texttt{create\_annotation\_file()} is
necessary; this file can be put in the \emph{Grid Search}, but it is not
mandatory.\\

Once the folder and the file are ready, run:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Grid\_search }\OtherTok{\textless{}{-}} \FunctionTok{perform\_grid\_evaluation}\NormalTok{(}
\NormalTok{  records,}
  \AttributeTok{sessions\_folder =} \StringTok{"Grid\_Search"}\NormalTok{,}
  \AttributeTok{prev\_classification =}\NormalTok{ records,}
  \DocumentationTok{\#\# Model parameters (can be changed by users)}
  \AttributeTok{resample =} \FunctionTok{c}\NormalTok{(}\ConstantTok{FALSE}\NormalTok{, }\ConstantTok{TRUE}\NormalTok{),}
  \AttributeTok{n\_init =} \FunctionTok{c}\NormalTok{(}\DecValTok{50}\NormalTok{, }\DecValTok{100}\NormalTok{, }\DecValTok{250}\NormalTok{, }\DecValTok{500}\NormalTok{),}
  \AttributeTok{n\_models =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{60}\NormalTok{),}
  \AttributeTok{pos\_mult =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{),}
  \AttributeTok{pred\_quants =} \FunctionTok{list}\NormalTok{(}
    \FunctionTok{c}\NormalTok{(.}\DecValTok{1}\NormalTok{, .}\DecValTok{5}\NormalTok{, .}\DecValTok{9}\NormalTok{),}
    \FunctionTok{c}\NormalTok{(.}\DecValTok{05}\NormalTok{, .}\DecValTok{5}\NormalTok{, .}\DecValTok{95}\NormalTok{),}
    \FunctionTok{c}\NormalTok{(.}\DecValTok{01}\NormalTok{, .}\DecValTok{5}\NormalTok{, .}\DecValTok{99}\NormalTok{)}
\NormalTok{  )}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

with \texttt{records} being the fully annotated data frame or a file
path. \texttt{prev\_classification} can be used to import a different
classification file with prelabeled records.

The hyperparameters are the following:

\begin{itemize}
\tightlist
\item
  resample: whether to use bootstrapping to increase variability between
  the ensemble models;
\item
  n\_init: number of initially labelled records to train the machine
  learning model on;
\item
  n\_models: the number of models to fit and then average to decrease
  the variance due to random sampling in the MCMC algorithm. It reduces
  the likelihood of extreme predictions;
\item
  pos\_mult: the oversampling rate of positive records. Useful to
  improve sensitivity;
\item
  pred\_quants: the distribution quantiles of each PPD, used to define
  the uncertainty zone and identify records to review.
\end{itemize}

Once the grid search is performed, the function
\texttt{analyse\_grid\_search()} can be used to analyse its results.
This function uses a partition tree analysis to identify hyperparameter
sets (\textbf{performance clusters}) with different mean performance
scores. As said before, the default score is Sensitivity x Efficiency;
alternatives are the Positive Rate (i.e., positive records found over
total records) or Positive Rate x Sensitivity.\\
The function produces a data frame with the score and the cluster of
reference of each combination of parameters, the best hyperparameter set
(i.e., the most efficient set among the most sensitive in the best
cluster) and the best set for each cluster. Also, the function creates a
plot to visualise each hyperparameter's impact on performance.

Once chosen, the hyperparameters can be passed to
\texttt{enrich\_annotation\_file()}.\\
Mind that the grid search process may take many days, even on powerful
computers, so we suggest sticking to the default set of parameters when
using \texttt{enrich\_annotation\_file()}.\\

\end{document}