### 2022-02-22 README.Rmd
added README and fixed S1 for the github page

---
title: "BART_SystReviewClassifier"
output: github_document
---

```{r readme_setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Description

This repo is the companion to the paper "An open-source integrated framework for the automation of citation collection and screening in systematic reviews" published as preprint at https://arxiv.org/abs/2202.10033. The following instructions are taken from the Supplemental Material S1.

```{r child = file.path("Manuscript", "S1.Protocol.Rmd")}
```