### 2022-02-22 Manuscript/Manuscript.Rmd
more language check and updated abstract

---
title: An open-source integrated framework for the automation of citation collection and screening in systematic reviews. 
authors:
  - name: Angelo D'Ambrosio
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: angelo.d.ambrosio@uniklinik-freiburg.de; a.dambrosioMD@gmail.com
  - name: Hajo Grundmann
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: hajo.grundmann@uniklinik-freiburg.de
  - name: Tjibbe Donker
    department: Institute for Infection Prevention and Hospital Hygiene
    affiliation: Freiburg University Hospital
    location: Freiburg, Germany
    email: tjibbe.donker@uniklinik-freiburg.de
abstract: |
  The exponential growth of scientific production makes secondary literature abridgements increasingly demanding.
  We introduce a new open-source framework for systematic reviews that significantly reduces time and workload for collecting and screening scientific literature.\
  The framework provides three main tools:
  1) an automatic citation search engine and manager that collects records from multiple online sources with a unified query syntax,
  2) a Bayesian, active machine learning, citation screening tool based on iterative human-machine interaction to increase predictive accuracy and,
  3) a semi-automatic, data-driven query generator to create new search queries from existing citation data sets.\
  To evaluate the automatic screener's performance, we estimated the median posterior sensitivity and efficiency [90% Credible Intervals] using Bayesian simulation to predict the distribution of undetected potentially relevant records.\
  Tested on an example topic, the framework collected 17,755 unique records through the citation manager;
  766 records required human evaluation while the rest were excluded by the automatic classifier;
  the theoretical efficiency was 95.6% [95.3%, 95.7%] with a sensitivity of 100% [93.5%, 100%].\
  A new search query was generated from the labelled dataset, and 82,579 additional records were collected;
  only 567 records required human review after automatic screening, and six additional positive matches were found.
  The overall expected sensitivity decreased to 97.3% [73.8%, 100%] while the efficiency increased to 98.6% [98.2%, 98.7%].\
  The framework can significantly reduce the workload required to conduct large literature reviews
  by simplifying citation collection and screening while demonstrating exceptional sensitivity.
  Such a tool can improve the standardization and repeatability of systematic reviews.
keywords:
  - Systematic review automation
  - Citation management
  - Online data collection
  - Active machine learning
  - Natural language processing
  - Bayesian modeling
bibliography: references.bib
lang: en-GB
output: 
  rticles::arxiv_article:
    number_sections: no
    keep_tex: true
    includes:
      in_header: header.tex
---

```{r, echo=FALSE, results='asis', message = FALSE, warning = FALSE, error = FALSE, dpi = 400, fig.width=12, fig.height=5}

out <- unlist(lapply(c(
  "Manuscript_Introduction.Rmd",
  "Manuscript_Methods.Rmd",
  "Manuscript_Results.Rmd",
  "Manuscript_Discussion.Rmd"
), knitr::knit_child, quiet = TRUE))

cat(out, sep = "\n")
```

\newpage

# References