### 2022-03-03 Manuscript/Manuscript.Rmd
prepare BMC submission

---
title: An open-source integrated framework for the automation of citation collection and screening in systematic reviews. 
authors:
  - first_name: Angelo
    second_name: D'Ambrosio
    initials: A.D.
    email: [angelo.d.ambrosio@uniklinik-freiburg.de, a.dambrosioMD@gmail.com]
    affiliation_IDs: aff1
    corresponding: true
  - first_name: Hajo
    second_name: Grundmann
    initials: H.G.
    email: hajo.grundmann@uniklinik-freiburg.de
    affiliation_IDs: aff1
  - first_name: Tjibbe
    second_name: Donker
    initials: T.D.
    email: tjibbe.donker@uniklinik-freiburg.de
    affiliation_IDs: aff1
affiliations:
  - id: aff1
    affiliation: Freiburg University Hospital
    department: Institute for Infection Prevention and Hospital Hygiene
    city: Freiburg
    country: Germany
abstract:
  - section_title: Background
    section_text: |
     The exponential growth of scientific production makes secondary literature abridgements increasingly demanding.
     We introduce a new open-source framework for systematic reviews that significantly reduces time and workload for collecting and screening scientific literature.
  - section_title: Methods
    section_text: |
     The framework provides three main tools:
     1) an automatic citation search engine and manager that collects records from multiple online sources with a unified query syntax,
     2) a Bayesian, active machine learning, citation screening tool based on iterative human-machine interaction to increase predictive accuracy and,
     3) a semi-automatic, data-driven query generator to create new search queries from existing citation data sets.\
     To evaluate the automatic screener's performance, we estimated the median posterior sensitivity and efficiency [90% Credible Intervals] using Bayesian simulation to predict the distribution of undetected potentially relevant records.\
  - section_title: Results
    section_text: |
     Tested on an example topic, the framework collected 17,755 unique records through the citation manager;
     766 records required human evaluation while the rest were excluded by the automatic classifier;
     the theoretical efficiency was 95.6% [95.3%, 95.7%] with a sensitivity of 100% [93.5%, 100%].\
     A new search query was generated from the labelled dataset, and 82,579 additional records were collected;
     only 567 records required human review after automatic screening, and six additional positive matches were found.
     The overall expected sensitivity decreased to 97.3% [73.8%, 100%] while the efficiency increased to 98.6% [98.2%, 98.7%].\
  - section_title: Conclusions
    section_text: |
     The framework can significantly reduce the workload required to conduct large literature reviews
     by simplifying citation collection and screening while demonstrating exceptional sensitivity.
     Such a tool can improve the standardization and repeatability of systematic reviews.
keywords:
  - Systematic review automation
  - Citation management
  - Online data collection
  - Active machine learning
  - Natural language processing
  - Bayesian modeling
abbreviations:
  - abbr: API
    descr: Application Programming Interface
  - abbr: BART
    descr: Bayesian Additive Regression Trees
  - abbr: COVID-19
    descr: Coronavirus Disease 2019
  - abbr: CR
    descr: Classification & Review
  - abbr: DTM
    descr: Document-Term Matrix
  - abbr: ICASR
    descr: International Collaboration for the Automation of Systematic Reviews
  - abbr: IEEE
    descr: Institute of Electrical and Electronics Engineers
  - abbr: IS
    descr: Inclusion Stability
  - abbr: MCMC
    descr: Monte Carlo Markov Chain
  - abbr: MeSH
    descr: Medical Subject Headings
  - abbr: MCMC
    descr: Monte Carlo Markov Chain
  - abbr: NLP
    descr: Natural Language Processing
  - abbr: PPD
    descr: Expected Value Posterior Predictive Distribution
  - abbr: RR
    descr: Relative Risk
  - abbr: SRA
    descr: Systematic Review Automation
  - abbr: SRH
    descr: Systematic Review Helpers
  - abbr: TF-IDF
    descr: Term Frequency - Inverse Document Frequency
  - abbr: WOS
    descr: Web Of Science
  - abbr: WSoR
    descr: Work Saved over Random
bibfile: references.bib
nocite: |
 @d2022open, @zenododata
appendix:
  acknowledgments: |
   We would like to thank Deborah Lawrie-Blum and Fabian Bürkin respectively for
   English language proofreading and mathematical formalization proof check.
  funding: |
   This project was developed under the Joint Programming Initiative on
   Antimicrobial Resistance (JPIAMR) through the 7th call, project number
   01KI1831 funded by BMBF and administrated by DLR Project Management Agency.
  data_availability: |
   The code and the instructions necessary to reproduce the results are
   available at https://github.com/AD-Papers-Material/BART_SystReviewClassifier.
   An updated, ready-to-use version of the framework is available at
   https://github.com/bakaburg1/BaySREn. All relevant data necessary to
   replicate the results is available at:
   https://doi.org/10.5281/zenodo.6323360.
  ethic_approval: Not applicable.
  competing_interest: The authors declare no competing interests.
  publication_consent: Not applicable.
  author_contribution: |
   Conceptualization: AD, TD; Data curation: AD;
   Formal analysis & Methodology: AD; Project administration: AD, TD;
   Software development: AD; Supervision: TD, HG;
   Writing–original draft: AD, TD; Writing–review & editing: AD, TD, HG.
   All authors have read and approved the manuscript.
  author_information: false
  additional_files:
    - title: S1. Framework description and usage.
      description: |
       Instruction on how to use the framework and
       reproduce the results in the manuscript.
    - title: S2. Additional outputs.
      description: Additional analysis outputs described in the manuscript.
    - title: S3. Online search instructions.
      description: |
       Instructions about how to prepare the framework to interact with the
       online scientific databases to collect records.
lang: en-GB
double_spacing: false
line_numbers: false
two_columns: false
remove_fig_numbering: true
output: 
  pdf_document:
    template: bmc_template.tex
    md_extensions: -auto_identifiers-autolink_bare_uris
    keep_tex: true
    keep_md: true
    pandoc_args:
      - --natbib
      - --lua-filter=bmc_filters.lua
    number_sections: false
    includes:
      in_header: header.tex
---

<!-- Background -->
```{r, child=c("Manuscript_Introduction.Rmd")}
```

<!-- Methods -->

```{r, child=c("Manuscript_Methods.Rmd")}
```

<!-- Results -->

```{r, child=c("Manuscript_Results.Rmd")}
```

<!-- Discussion -->

```{r, child=c("Manuscript_Discussion.Rmd")}
```